% The Acknowledgements page, for thanking everyone
\clearpage
\setstretch{1.3} % Reset the line-spacing to 1.3 for body text (if changed)
\acknowledgements{
    \addtocontents{toc}{\vspace{1em}} %Add a gap in the Contents, for aesthetic
        
        I would like to take this opportunity to recognise, and commemorate in print, my thanks to all those who have helped me throughout my studies.
    
I would like to start by thanking the support and guidance of my supervisors Professors Chris Nugent and Sally McClean. Chris for the \textit{phenomenal} amount of support, guidance and friendship during my time at Ulster. Chris has overwhelmed me with great opportunities, and supported me wholeheartedly through them, for which I am forever grateful. Sally has been invaluable to my research, with her vast knowledge, a keen eye for detail and a wit to match, it has been a pleasure to work with her. I would like to extend my thanks to all those involved in the Gray Matters and TAUT projects. In particular Professor Maria Norton for her contagious motivation, guidance and expertise and also Dr Ian Cleland for his guidance in my day-to-day research, always with a positive outlook and great sense of humour.
 
To my family, I want to thank you all for your love and support in everything that I do. To my father, Graeme, I thank you for your dependable guidance, rock solid advice, and for always being just a phone-call away. To my mother, Sharon, I want to thank you for your literary mind, your confidence in me and for pushing me when I was young, I realise now what it was all for. To my sister, Natalie, we are one in the same, and we all know \textit{you're the smart one} and I am so very proud of you. Of course, I want to thank David and Dolores McCool for all their encouragement, humour and a never ending offerings of food over the years. I would also like to thank all my friends for their humbling indifference toward my work, and focusing our time purely on fun. You know who you all are.

Lastly, I would like to dedicate this thesis to my fiancee Ce\'ara McCool. Thank you for your unyielding support, encouragement, patience, and love. Your conscientious attitude to life has always motivated me to keep pushing forward. Ceara, the very fact you are reading this means this thesis is now complete, no more words left to write. This chapter of my life has come to an end\ldots I can't wait to start the next one together. 
}