\chap{Conclusion and Future Work}
\label{chapter: conclusion}

This Chapter concludes and summarises the results from thesis studies, summarising the findings and producing a number of messages for various research fields and practices.

\section{Thesis Summary}
Through this body of work the underlying theme has been to exploit the ubiquitous nature of smartphones to improve in health and wellbeing outcomes. This has been demonstrated in two distinct cases studies which span both the treatment and prevention areas of AD. 

Chapter \ref{chapter: treatment-framework} focused on using the ever available smartphone as the platform for an assistive reminder tool for PwD. The adoption of the solution from the cohort was encouraging, however, over-time exposed fundamental issues regarding technology adoption with the elderly and cognitively impaired. Despite a lack of widespread adoption at the end of the 12 months, the data from the study permitted the concept of using smartphone sensors to predict if a reminder will be acknowledged or missed, to be proven. The following question, yet to be answered, is if this will improve the acknowledgement rates of reminders, and if that will consequently result in improved adherence to the actual tasks being reminded. As such, the models described in the Chapter need be implemented and evaluated with another, comparable, cohort.

Chapters \ref{chapter: prevention-framework}, \ref{chapter: prevention-evaluation}, \ref{chapter: prevention-rctresults} were focused heavily on the areas of dementia prevention through behaviour change, facilitated by a smartphone app. The works demonstrated how disease risk could be mitigated by applying a process framework developed by the author, to a disease area. The result from this produced an app which guided and encouraged users, step by step, to change disease associated behaviours, resulting in continual health improvement. Unlike the work in Chapter \ref{chapter: treatment-framework}, the study participants had no such perceived barriers to adoption, and as such the case study truly exploited the pervasive nature of the smartphone, yielding every advantage offered to integrate with an individual’s daily life and routines. 
 
\subsection{Contributions to Knowledge}
Throughout the works detailed in this Thesis, a number of areas were identified by the author where contributions to knowledge have been made. These have been listed below, and briefly summarised in the following subsections:
\begin{itemize}
	\item Developed mobile-app behavioural change framework and guidelines.
	\item Demonstrated behavioural effects of applied framework.
	\item Demonstrated clinical efficacy of applied framework.
	\item Detailed impartial approach to critically evaluate and rate mHealth apps.
	\item Demonstrated use of sensor data to infer opportune moments to deliver reminders, for PwD and healthy individuals.
	\item Identified areas for future mHealth-supported behavioural research.
	\item Identified areas for future research of reminder tools for PwD.
\end{itemize}
%VIVA: Chris -These are all good!  Do they map easily on to the Chapters?
\subsubsection{Mobile-app Behavioural Change Framework and Guidelines}
The primary contribution to knowledge in this thesis is the formalisation of a an evidence based process to develop mobile-apps to encourage and monitor positive behaviour change. In addition, the framework has been made modular, in that any component can be applied to a specific health condition and scale accordingly. The framework was applied to the area of behaviour change for the purpose of AD risk reduction in middle aged adults, resulting in the Gray Matters app, which was the primary intervention component in a 6 month behaviour change RCT.  

\subsubsection{Behavioural Effects of Applied Framework}
The results of the RCT displayed that a resulting app from the author's framework had successfully managed to facilitate numerous stages of the TTM, namely preparing participants for change, allowing them to accurately monitor and asses their actions, and encourage continued maintenance and improvement of their desired behaviours. Positive results from the RCT exit survey also suggested that the majority of users wished to continue their behaviour change efforts, supported by the app. 

\subsubsection{Clinical Efficacy of Applied Framework}
As a result of sustained behaviour change through usage of the app, promising and statistically significant clinical effects were observed in the treatment cohort. Positive effects were observed in the cohorts BMI, cholesterol (HDL, LDL, and total), systolic and diastolic blood pressures, and resting heart rate. A high frequency of app usage (7+ times per week) was found to result in the largest reduction in BMI and blood pressure, demonstrating the framework's  ability to engage and retain engagement. Participants who utilised and achieved over 60\% of their recommended daily goals were also 1.762 times more likely to decrease their BMI during the study. This also highlights the frameworks ability in establishing suitable and attainable user goals to combat the targeted condition. 

\subsubsection{Impartial Approach to Evaluate and Rate mHealth Apps}
The author details in \ref{chapter: prevention-evaluation} the results of a expert review of the developed Gray Matters app, including methods to ensure an imparial process. In addition, the author performs content analysis on 87 existing mHealth apps, contrasts with the reviewers scores and uncovered an additional 18 observations, which included a number of significant recommendations to be applied to the author's original framework.


\subsubsection{Infer Opportune Reminder Moments Using Sensor Data}
The author demonstrated the use of embedded smartphone sensors to infer opportune moments to deliver reminders based upon sensor data collected from PwD and healthy individuals over a 12-month period. The works detailed the optimal features for each sensor, and demonstrated merit of of ensemble classifiers using decision trees to boost classification performance over previously detailed approaches. 


\subsection{Future Work}
Throughout the works, during reviews of existing literature, analysis of study data, and interviews with study participants, a number of areas were identified by the author in which future study could improve, or elicit new and interesting, findings.

\subsubsection{Advanced Personalisation and Feedback}
Interviews and questionnaires performed by the participants in the Gray Matters study indicated that \textit{truly} personalised feedback was greatly desired. Specifically, participants wished to view personalised progress reports, based upon their own efforts, rather than those imposed by global recommendations. Whilst the study collaborators had foreseen this, it was agreed that it would detract from assessing motivation for established goals. In retrospect however, it may have been possible to facilitate both, as discussed in Section \ref{subsection: gm-future-personalisation}. As the demand for such a feature was so high, it is anticipated that personalised goal setting may significantly increase both adoption and user retention, two key factors for successful behaviour change.

\subsubsection{Seek External Quantifiable Data Sources}
In the Gray Matters study, the validity of the users reported behaviours was dependant on their honesty and ability to accuracy assess their own behaviours. As with the wearable activity monitor in the physical domain, it may be possible to supplement user reported behaviours in other domains with quantifiable data recorded from external sources, such as additional wearable sensors, sleep monitors, observing gaming activity, or data-mining from social networking. The results of such an approach could yield greater insight into the accuracy of self-reporting, and serve to correct reported values. Quantifiable data observed from a third party is also enlightening to the user, and can serve to correct any bias about their own behaviours \cite{Elliott2014}.

\subsubsection{Artificial Intelligence Coaching}
In the Gray Matters study, participants had reported that there were various times when a fact and suggestion pair was not fully understood, or had raised additional personal queries, for which the app had no answers. Often the participants would try to reach out to the study investigators, in some cases eventually turning to websites for answers. The queries were simple, and typically related to diet and exercise. An example scenario:
Daily suggestion presents: 'Eat 2 cups of brown rice instead of chips with dinner tonight'. The participant would like to know if white rice is OK, as they do not own brown rice. 
It is a simple query, and the answer is relatively inconsequential, however, a lack of answers extrapolated over numerous users and time, will result in loss of engagement, and thus loss of impact. A human operated query answering facility to service a large user based is also not feasible. Nevertheless, great progress is currently being made in natural language engineering, enabling the ability to answer open questions through databases of expert knowledge \cite{HIRSCHMAN2001, Fader2014}. Application of these techniques, whilst advanced, will allow researchers to develop a deeper suggestion base, and provide a much higher level of support to end-users.
 
\subsubsection{Evaluation of sensorised reminder delivery system}
The work performed in Chapter \ref{chapter: treatment-framework} needs to be extended to fully achieve the original goal of improving adherence to reminders for PwD. The current work provides a solid technological base from which the models may be implemented on a smartphone, or cloud. The app was written in a modular approach, with the intension to be extended by the author and others in the field. Currently, the base application has been extended by \citeauthor{Patterson2015} who aim to use the platform to assess and eventually improve task completion followthrough, once a reminder has been issued \cite{Patterson2015}.

\subsubsection{Carer-focused Apps}
Carers of PwD are often family members, whose age would be similar as the middle-aged cohort in the Gray Matters study, who posed very minimal barriers to successful technology adoption. Given the opportunity to expand upon the work performed in Chapter \ref{chapter: treatment-framework} the author would desire to create a carer orientated solution, removing the need for the PwD to have any interaction with technology.

\section{Messages for Behavioural Scientists}
\textbf{Harness the Smartphone.}
Changing behaviours may be the key to tackling the leading causes of mortality and morbidity, such as poor diet choices, lack of exercise, smoking and alcohol consumption. Diet and exercise apps to quantify efforts do exist, some with their foundations in exercise and nutritional science, however, there is a gaping void where behavioural and psychological science based apps should be. 
The smartphone, like no other tool before it, has become part of everyday life, embedded into the behaviours of millions, giving direct access to quantifiable data. Through the understanding of behaviours on a global scale, the development of all health interventions will improve. It is the author's opinion, that the merging of behavioural science with ubiquitous computing, currently in the form of smartphones, stands to empower the individual, and revolutionise healthcare on a global scale. 

\section{Message for Technologists}
\textbf{Technology is the question, not the answer.}
Whilst technology may hold some of the answers, it is not \textit{the} answer. Before applying technology to a problem area, ask, \textit{does it really belong there?}
A great level of diligent research must be performed, drawing knowledge from every available perspective, through collaboration and interdisciplinary research. Whilst it is true that technology-led health-orientated projects often advance the state-of-the-art, it seems seldom that they actually prove efficacy over more traditional, non-technology based approaches. With increased focus on efficacy and impact, funding and reputations are at stake. It is up to us, the technologists, to prove that the field deserves the spotlight that it gets, through true collaboration, not discordant co-operation.